---
title: "PS 1 SOLUTION SET"
author: Spring 2024
output: 
  pdf_document:
    includes:
      in_header: header.tex
---


There are usually many ways to arrive at complete and correct answers. The solution set shows you one way. As always, study the keys (the `.Rmd` code and `.pdf` output), and reach out with questions. 

```{r setup, include=TRUE}
## Set code chunk options
  knitr::opts_chunk$set(
    echo = TRUE,
    eval = TRUE,
    fig.align = 'center',
    message = FALSE,
    warning = FALSE
  )
```

```{r}
## Load packages
  library(tidyverse)
  library(knitr)

## PS1 data
  df = gapminder::gapminder_unfiltered
```

1. The unfiltered Gapminder data is a country-year dataset with `r nrow(df)` observationsa and `r length(df)` variables. The data covers `r df |> pull(country) |> unique() |> length()` unique countries from `r min(df['year'])` to `r max(df['year'])`.

2. The plot below scatters life expectancy over year for the full dataset.

```{r plot1, fig.dim=c(3,2), fig.cap = 'Life expectancy over time'}
ggplot(df, aes(x = year, y = lifeExp)) +
  geom_jitter(
    shape = 21,
    alpha = 0.3
  ) +
  geom_smooth(
    color = 'firebrick',
    se = F
  ) +
  labs(
    x = 'Year',
    y = 'Life Expectancy'
  ) +
  theme_classic()
```

3. The table below summarizes coverage of the Gapminder (unfiltered) data by continent. 

```{r tab}
df |>
  count(continent, name = 'Freq') |>
  mutate(
    Per = Freq / sum(Freq) * 100
  ) |>
  kable(
    digits = 1L, 
    format.args = list(big.mark = ','),
    caption = 'Gapminder Coverage by Continent'
  )
```

4. See the recreated plot using the unfiltered data below. 

```{r recplot, fig.dim=c(4,3), fig.cap = 'GDP over time by continent'}
  df |>
    filter(continent != "Oceania") |>
    ggplot(aes(x = gdpPercap, y = lifeExp)) +
  	facet_wrap(~ continent) +
  	geom_point(
  	  alpha = (1/3), 
  	  size = 3
  	) +
    geom_smooth(
  	  lwd = 1.5, 
  	  se = FALSE
  	) +
  	scale_x_log10() # log the x axis the right way

```

<!-- Howdy! I'm extra credit 2. If you find me, email Prof Hart. -->

5. See the styled code below. Note especially the spacing and use of line breaks to organize code within chunks and between functions/operator.

```{r, eval = FALSE}
  flights |>
    filter(dest == "IAH") |>
    group_by(
      year, 
      month, 
      day
    ) |>
    summarize(
      n = n(), 
      delay = mean(arr_delay, na.rm = TRUE)
    ) |>
    filter(n > 10)

flights |>
  filter(
    carrier == "UA", 
    dest %in% c("IAH", "HOU"), 
    sched_dep_time > 0900,
    sched_arr_time < 2000
  ) |>
  group_by(flight) |>
  summarize(
    delay = mean(arr_delay, na.rm=TRUE), 
    cancelled = sum(is.na(arr_delay)), 
    n = n()
  ) |>
  filter(n > 10)
```
